Pharmaceutical Supply Chain Forecasting by using Machine Learning in Ethiopian Pharmaceutical Supply Services

### 2024-10-09 manuscript/02-literature.qmd
Revised sections

# Research background {#sec-lit}
# focus on forecasting on pharmaceuticals products 
# 2 pages 
# how forecasting is done 
# what other researches had done 
# what are the limitation of the other papers
# 15 relevant papaers

# 02- literature

@@

---
title: "Table 1. Summary of Some Studies on Forecasting in pharmaceuticals products"
tbl-cap-location:top

```{r}

library(knitr)
data <- data.frame(
  Header1 = c("Row 1", "Row 2", "Row 3"),
  Header2 = c("Data 1", "Data 2", "Data 3"),
  Header3 = c("Data 1", "Data 2", "Data 3")
)
kable(data, format = "html")

```

 |                       |           |               |             |          | References       |  | Year      | Variable      |             |
 | :-----------           |-----------| ------------- |-----------  |
 | Siddiqui et al            |           |               |             | 
 |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               | |
 |            |           |               |
 |            |           |               |         
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |               |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |          
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           ||            |           |
 | -----------|-----------|
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |          
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |          
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |           |
 |            |
 |            |
 |            |
 |            |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |
 |

>>>>

# Research background {#sec-lit}

@@

### 2024-10-09 manuscript/03-experiment.qmd
Revised sections

# you need to talk about the data set 
# some visualization 
# make reduce to paragaphs minimize it 
# forecast accuracy measures provid ethe formulas

## Data {#sec-data}
# 2.1. Dataset
For this study, we utilized a dataset spanning five years (2018-2022) of sales data obtained from EPSS. Rigorous checks were conducted to ensure the consistency and completeness of the collected data. From the extensive pool of pharmaceutical products within EPSS, we strategically selected a set of 37 key pharmaceuticals, encompassing various programs and representing different classes of drugs. This deliberate selection enriches the research by providing a comprehensive representation across the pharmaceutical landscape.
We are presenting the graph for some of the pharmaceuticals sales for the last five years by using visualization for the readers to undesrtand the sales of the pharmaceuticals.
```{r}
#| include: false
library(tidyverse)
library(lubridate)
library(fpp3)
library(ggthemes)
monthly_issues_tsibble <- read_rds("../data/monthly_issues_tsibble.rds")
monthly_issues_tsibble <- monthly_issues_tsibble |> filter_index("2017 Aug" ~ "2022 Jun")

@@

# Visualising monthly time series

## Title

>>>>

## Data {#sec-data}

### Dataset

For this study, we utilized a dataset spanning five years (2018-2022) of sales data obtained from EPSS. Rigorous checks were conducted to ensure the consistency and completeness of the collected data. From the extensive pool of pharmaceutical products within EPSS, we strategically selected a set of 37 key pharmaceuticals, encompassing various programs and representing different classes of drugs. This deliberate selection enriches the research by providing a comprehensive representation across the pharmaceutical landscape.
We are presenting the graph for some of the pharmaceuticals sales for the last five years by using visualization for the readers to undesrtand the sales of the pharmaceuticals.
```{r}
#| include: false
library(tidyverse)
library(lubridate)
library(fpp3)
library(ggthemes)
monthly_issues_tsibble <- read_rds("../data/monthly_issues_tsibble.rds")
monthly_issues_tsibble <- monthly_issues_tsibble |> filter_index("2017 Aug" ~ "2022 Jun")

@@

## Visualising monthly time series

### 2024-10-09 manuscript/04-results.qmd
Revised sections

# Results and discussion {#sec-results}
```{r}
#| include: false
#| label: restults
library(readr)
fc_accuracy_overall__without_predictor <- read_csv("../result/fc_accuracy_overall__without_predictor.csv")
fc_accuracy_predictor_overall <- read_csv("../result/fc_accuracy_predictor_overall.csv")
fc_accuracy_overall <- read_csv("../result/fc_accuracy_all_item.csv")
data_visualisation <- read.csv("../rscript/data_visualisation.R")
```
from the total of 37 key pharmaceuticals selected and tested with the  different forecasting models arima showed that a better forecast accuracy for most of the pharmaceuticals followed by mean and  regresstion models with RMSSE value of 0.850, 0.942 and 0.945 respectively (Table 1).

```{r}
library(kableExtra)
library(dplyr)  # Load the dplyr package
fc_accuracy_cleaned <- na.omit(fc_accuracy_overall__without_predictor)
fc_accuracy_table <- fc_accuracy_cleaned %>%
  dplyr::select(.model, RMSE, MAE, MASE, RMSSE, winkler, CRPS)  # Explicitly specify dplyr::select
kbl(fc_accuracy_table, booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), full_width = FALSE)
```
Then we tested the models by adding different predictors which affects the sales of the pharmaceuticals like physical count, budget release and budget closure by the government, COVID 19, campaign in to the models as predictors which will affects the sales of the key pharmaceuticals products and the result showed that still arima have a better forecast accuracy with RMSSE of 0.850 (Table 2). 
```{r}
#| label: result3
#| tbl-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
fc_accuracy_predictor_overall %>% 
  kbl(booktabs = TRUE, escape = FALSE) |>
  kable_styling(latex_options = c("hold_position"),full_width = F)
```
```{r}
#| label: result2
#| fig-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
library(ggplot2)  # Load the ggplot2 package
library(forcats)  # Load the forcats package
fc_accuracy_overall %>%
  select(`.model`, item, RMSSE) %>% 
  ggplot(mapping = aes(x = RMSSE, y = fct_reorder(`.model`, RMSSE, .fun = median))) +
  geom_boxplot()
```
Individual item with different forecasting models 
We tested 10 models namely arima, arima with predictor, combination, exponential smoothing, mean, naive, regression, regression with predictor, snaive and stlf. Individually on the selected key pharmaceuticals and the result indicated that 11 (32.43%) of the selected pharmaceuticals have better forecast accuracy for arima predictor models, followed by naive 6 (16.21%), mean, regresstion, arima, and STLF methods each of them with 3 (8.10%) (Fig 2)

@@

```
Those pharmaceuticals who had a better forecast accuracy with Arima predictors are Amoxacillin 500mg capsule, Anti RhD, Artehmether + Lumefantrine, Atenolol, Atrovastatin 200mg, Ferrosu sulphate with follic acid, frusemide injection, Insuline Iosphane bipahsic, Lamivudine +Efaverienze + Tenofovir, Lidocaine + HCL and Propyl thiro uracil.The  six pharmaceuticasl which showed a better forecast accuracy with naive models were Ciprofloxacillin 500mg tablet, Magnesoum sulphate injection, Omeprazole capsule, Omeprazole injection, Rapid diagnostic test, Sulphamethaxone + Trimetoprine tablet.

>>>>

# Results and discussion {#sec-results}

```{r}
#| include: false
#| label: restults
library(readr)
fc_accuracy_overall__without_predictor <- read_csv("../result/fc_accuracy_overall__without_predictor.csv")
fc_accuracy_predictor_overall <- read_csv("../result/fc_accuracy_predictor_overall.csv")
fc_accuracy_overall <- read_csv("../result/fc_accuracy_all_item.csv")
data_visualisation <- read.csv("../rscript/data_visualisation.R")
```

from the total of 37 key pharmaceuticals selected and tested with the  different forecasting models arima showed that a better forecast accuracy for most of the pharmaceuticals followed by mean and  regresstion models with RMSSE value of 0.850, 0.942 and 0.945 respectively (Table 1).

```{r}
library(kableExtra)
library(dplyr)  # Load the dplyr package
fc_accuracy_cleaned <- na.omit(fc_accuracy_overall__without_predictor)
fc_accuracy_table <- fc_accuracy_cleaned %>%
  dplyr::select(.model, RMSE, MAE, MASE, RMSSE, winkler, CRPS)  # Explicitly specify dplyr::select
kbl(fc_accuracy_table, booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), full_width = FALSE)
```

Then we tested the models by adding different predictors which affects the sales of the pharmaceuticals like physical count, budget release and budget closure by the government, COVID 19, campaign in to the models as predictors which will affects the sales of the key pharmaceuticals products and the result showed that still arima have a better forecast accuracy with RMSSE of 0.850 (Table 2).

```{r}
#| label: result3
#| tbl-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
fc_accuracy_predictor_overall %>% 
  kbl(booktabs = TRUE, escape = FALSE) |>
  kable_styling(latex_options = c("hold_position"),full_width = F)
```
```{r}
#| label: result2
#| fig-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
library(ggplot2)  # Load the ggplot2 package
library(forcats)  # Load the forcats package
fc_accuracy_overall %>%
  select(`.model`, item, RMSSE) %>% 
  ggplot(mapping = aes(x = RMSSE, y = fct_reorder(`.model`, RMSSE, .fun = median))) +
  geom_boxplot()
```

Individual item with different forecasting models 
We tested 10 models namely arima, arima with predictor, combination, exponential smoothing, mean, naive, regression, regression with predictor, snaive and stlf. Individually on the selected key pharmaceuticals and the result indicated that 11 (32.43%) of the selected pharmaceuticals have better forecast accuracy for arima predictor models, followed by naive 6 (16.21%), mean, regresstion, arima, and STLF methods each of them with 3 (8.10%) (Fig 2)

@@

```

Those pharmaceuticals who had a better forecast accuracy with Arima predictors are Amoxacillin 500mg capsule, Anti RhD, Artehmether + Lumefantrine, Atenolol, Atrovastatin 200mg, Ferrosu sulphate with follic acid, frusemide injection, Insuline Iosphane bipahsic, Lamivudine +Efaverienze + Tenofovir, Lidocaine + HCL and Propyl thiro uracil.The  six pharmaceuticasl which showed a better forecast accuracy with naive models were Ciprofloxacillin 500mg tablet, Magnesoum sulphate injection, Omeprazole capsule, Omeprazole injection, Rapid diagnostic test, Sulphamethaxone + Trimetoprine tablet.
